### 2019-04-05 0_abstract.tex
Small tweaks and edits.
Small tweak.
Tweak abstract.
Fiddle with spacing and text to make things fit again.

\begin{abstract}
% 126 words
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance envelopes of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench seeks to provide the ease of implementation and portability of
microbenchmarks, which makes comparisons between many systems
possible, with the ability to model a wide range of application
scenarios, giving it coverage approximating large benchmarks.

We conduct a comprehensive study with implementations of Task Bench in
13 programming systems on up to 256 Haswell nodes of the Cori
supercomputer. We introduce and use a novel metric, \emph{minimum effective task
  granularity} to study the baseline runtime overhead of each system
and the impact of different dependence patterns. We also
study each system's scalability, and ability to hide communication and
mitigate load imbalance.
\end{abstract}

>>>>

\begin{abstract}
% 128 words
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance envelopes of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench seeks to provide the ease of implementation and portability of
microbenchmarks, which makes comparisons between many systems
possible, with the ability to model a wide range of application
scenarios, giving it coverage approximating mini-apps.

We conduct a comprehensive study with implementations of Task Bench in
13 programming systems on up to 256 Haswell nodes of the Cori
supercomputer. We introduce and use a novel metric, \emph{minimum effective task
  granularity} to study the baseline runtime overhead of each system
and the impact of different dependence patterns. We also
study each system's scalability, ability to hide communication and
mitigate load imbalance.
\end{abstract}

### 2019-04-05 2_metg.tex
Fiddle with spacing and text to make things fit again.

The \emph{minimum effective task granularity}, METG(50\%), for an application $A$ is
the smallest average task granularity (i.e., task duration) such that $A$
achieves overall efficiency exceeding 50\%. For example, in
a compute-bound application efficiency might be measured as the
percentage of the available FLOPS achieved. On a machine with a peak
capability of 1.26 TFLOPS per node, METG(50\%) would correspond to
the smallest task granularity achieved while maintaining at least 0.63
TFLOPS. Absolute (rather than relative) efficiency is preferred
because it makes it easier to conduct fair comparisons with
programming systems that reserve one or more cores for internal use.

Figures~\ref{fig:flops-mpi} and \ref{fig:efficiency-mpi} show how METG is
calculated. The application, in this case an MPI implementation of the
Task Bench stencil pattern in Figure~\ref{fig:task-graphs-stencil}, is
run on a single Haswell node of Cori~\cite{Cori} with a problem size large enough to
ensure that it achieves peak FLOPS. This confirms
that the application is properly configured and that the
efficiency metric is achievable. The problem
size is then repeatedly reduced while maintaining exactly the same hardware
configuration (in particular, the same number of nodes). The
expectation is that as problem size shrinks,
performance will begin to drop and eventually approach zero, as shown in Figure~\ref{fig:flops-mpi}. Systems
with lower runtime overheads maintain higher performance at smaller
problem sizes compared to systems with higher overheads.

@@

In Figure~\ref{fig:efficiency-mpi}, we see that
efficiency starts near 100\%, and that initially task granularity
drops quickly with minimal change in efficiency. However as task
granularity shrinks further, efficiency begins to drop more rapidly, leading
eventually to asymptote. Intuitively, this asymptote
represents the fundamental baseline overhead of the system, the point
below which the system simply cannot execute tasks efficiently.

>>>>

The \emph{minimum effective task granularity}, METG(50\%), for an application $A$ is
the smallest average task granularity (i.e., task duration) such that $A$
achieves overall efficiency of at least 50\%. For example, in
compute-bound applications efficiency can be measured as the
percentage of the available FLOPS achieved. On Cori with 1.26 TFLOPS available per Haswell node, METG(50\%) corresponds to
the smallest task granularity achieved while maintaining at least 0.63
TFLOPS. Absolute (rather than relative) efficiency is preferred
because it makes it easier to conduct fair comparisons with
programming systems that reserve one or more cores for internal use.

Figures~\ref{fig:flops-mpi} and \ref{fig:efficiency-mpi} show how METG is
calculated. The application, in this case an MPI implementation of the
Task Bench stencil pattern in Figure~\ref{fig:task-graphs-stencil}, is
run on a single Haswell node of Cori with a problem size large enough to
ensure that it achieves peak FLOPS. This confirms
that the application is properly configured and that the
efficiency metric is achievable. The problem
size is then repeatedly reduced while maintaining exactly the same hardware
configuration (in particular, the same number of nodes). The
expectation is that as problem size shrinks,
performance will begin to drop and eventually approach zero, as shown in Figure~\ref{fig:flops-mpi}. Systems
with lower runtime overheads maintain higher performance at smaller
problem sizes compared to systems with higher overheads.

@@

In Figure~\ref{fig:efficiency-mpi}, we see that
efficiency starts at 100\%. Initially task granularity
shrinks with minimal change in efficiency. However as task
granularity shrinks further, efficiency begins to drop more rapidly, leading
eventually to asymptote. Intuitively, this asymptote
represents the fundamental baseline overhead of the system.

### 2019-04-05 4_implementation.tex
Fiddle with spacing and text to make things fit again.

There are also aspects of some systems that we do not consider in our
evaluation. For example, a number of systems provide some level of
support for heterogeneous processors such as GPUs, but the support
varies so widely (if it exists at all) that it seems difficult to
compare systems with direct GPU support to those wihtout in an
apples-to-apples manner.

@@

Swift/T~\cite{Wozniak13} is a parallel scripting language intended
primarily for the composition of HPC workflows. Tasks in Swift/T can be written in any
programming language, and may even be themselves parallel. Swift/T
programs follow dataflow semantics, where every statement may
potentially execute in parallel as soon as its dependencies are
satisfied; statements only execute sequentially when explicitly
requested. The Swift/T compiler performs a number of optimizations to
improve performance of highly parallel programs~\cite{Armstrong14}.

>>>>

There are also aspects of some systems that we do not consider in our
evaluation. For example, a number of systems provide some level of
support for heterogeneous processors such as GPUs, but the support
varies so widely (if it exists at all) that it seems difficult to
compare systems with direct GPU support to those without in an
apples-to-apples manner.

@@

Swift/T~\cite{Wozniak13} is a parallel scripting language intended
primarily for the composition of HPC workflows. Tasks in Swift/T can be written in any
programming language. Swift/T
programs follow dataflow semantics, where every statement may
potentially execute in parallel as soon as its dependencies are
satisfied; statements only execute sequentially when explicitly
requested. The Swift/T compiler performs a number of optimizations to
improve performance of highly parallel programs~\cite{Armstrong14}.

### 2019-04-05 5_evaluation.tex
Fiddle with spacing and text to make things fit again.

We present a comprehensive evaluation of our Task Bench implementations on up to 256
Haswell nodes of the Cori supercomputer~\cite{Cori}, a Cray XC40
machine. Cori Haswell nodes have 2 sockets with Intel Xeon E5-2698 v3
processors (a total of 32 physical cores per node), 128 GB RAM, and a
Cray Aries interconnect. We use GCC 7.3.0 for compiling all Task Bench
implementations, and (where applicable) the system default MPI
implementation, Cray MPICH 7.7.3. Additional versions and flags for the
various systems are shown in Table~\ref{tab:flags}.

@@

To answer this question, we use the METG metric to determine the
smallest task granularity that can be executed at a given level of
efficiency as a proxy for overhead. Figure~\ref{fig:metg-compute}
shows how METG(50\%) varies with node count for a variety of
dependence patterns supported by Task Bench. METG(50\%) is calculated
separately at each node count, to isolate runtime system behavior from
changes in communication latencies and topology when using
progressively larger portions of the machine. We consider the
following configurations of Task Bench:

\begin{itemize}
\item Figure~\ref{fig:metg-compute-stencil} is a 1D stencil where each
  task depends on 3 other tasks (including the same point in the
  previous timestep).
\item Figure~\ref{fig:metg-compute-nearest} is a pattern where each
  task depends on 5 others, chosen to be as close as possible.
\item Figure~\ref{fig:metg-compute-spread} is a pattern where each
  task depends on 5 others, spread as widely as possible.
\item Figure~\ref{fig:metg-compute-4x-nearest} shows 4 identical
  copies of the nearest dependence pattern executing concurrently.
\end{itemize}

We observe in the results that the baseline overheads of different
systems vary by over 5 orders of magnitude. It is worth
remembering when considering this metric that this is a \emph{minimum}
effective task granularity. Therefore applications with an average
task granularity of \emph{at least} this value can usually be expected
to execute efficiently. Typical average task granularity will
generally be determined by the application domain being
considered. Most notably, for data analytics workloads being run in
industrial data centers, the higher METG values observed for Spark and
TensorFlow are sufficient. In contrast, for high-performance
scientific simulations, task granularities in the millisecond range
are useful, as such applications communicate (e.g., for halo
exchanges) much more frequently.

In general, the least complicated pattern (stencil) is most favorable
to MPI, as it represents a truly bulk synchronous task graph with no
opportunity for task parallelism. For the stencil pattern, the
dominating factor is the basic overhead of executing a task, which is
minimal for MPI as the Task Bench implementation is bulk synchronous
and simply executes tasks one after another in alternation with
communication phases. In contrast, the asynchronous execution
mechanisms of the other systems are pure overhead in this scenario.

The gap between MPI and other systems shrinks as the complexity of the
communication pattern grows, and even reverses as task parallelism is
added in the form of multiple task graphs. Some of the slower systems
are omitted from comparisons with more complicated dependence
patterns, as they failed to complete in a reasonable time.

@@

The difference between the remaining systems is due primarily to the
scheduling behavior of the various systems. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Therefore, systems that provide an
additional on-node work stealing capability see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at very small task
granularities. This can be seen in the difference between Chapel
distrib (which uses a work stealing scheduler) vs Chapel with its
default (non-work-stealing) scheduler.

@@

Also of interest is the ability of asynchronous systems to
hide communication latency in the presence of task
parallelism. Figure~\ref{fig:efficiency-communication} plots task
granularity vs efficiency curves for varying amounts of
communication. The degree of communication is configured by the number
of bytes produced by each task (and therefore communicated with each
task dependency).

>>>>

We present a comprehensive evaluation of our Task Bench implementations on up to 256
Haswell nodes of the Cori supercomputer~\cite{Cori}, a Cray XC40
machine. Cori Haswell nodes have 2 sockets with Intel Xeon E5-2698 v3
processors (a total of 32 physical cores per node), 128 GB RAM, and a
Cray Aries interconnect. We use GCC 7.3.0 for all Task Bench
implementations, and (where applicable) the system default MPI
implementation, Cray MPICH 7.7.3. Versions and flags for the
various systems are shown in Table~\ref{tab:flags}.

@@

To answer this question, we use the METG metric to determine the
smallest task granularity that can be executed at a given level of
efficiency as a proxy for overhead. Figure~\ref{fig:metg-compute}
shows how METG(50\%) varies with node count for a variety of
dependence patterns supported by Task Bench. METG(50\%) is calculated
separately at each node count, to isolate runtime system behavior from
changes in communication latencies and topology when using
progressively larger portions of the machine.

We consider the following configurations of Task Bench:
Figure~\ref{fig:metg-compute-stencil} is a 1D stencil where each task
depends on 3 other tasks (including the same point in the previous
timestep). Figure~\ref{fig:metg-compute-nearest} is a pattern where
each task depends on 5 others, chosen to be as close as
possible. Figure~\ref{fig:metg-compute-spread} is a pattern where each
task depends on 5 others, spread as widely as possible. And
Figure~\ref{fig:metg-compute-4x-nearest} shows 4 identical copies of
the nearest dependence pattern executing concurrently.

We observe in the results that the baseline overheads of different
systems vary by over 5 orders of magnitude. It is worth
remembering when considering this metric that this is a \emph{minimum}
effective task granularity. Therefore applications with an average
task granularity of \emph{at least} this value can usually be expected
to execute efficiently. Typical task granularities will
generally be determined by the application domain being
considered. Most notably, for large-scale data analytics workloads, the higher METG values observed for Spark and
TensorFlow are sufficient. In contrast, for high-performance
scientific simulations, task granularities in the millisecond range
are useful, as such applications communicate (e.g., for halo
exchanges) much more frequently.

The least complicated pattern (stencil) is most favorable
to MPI, as it represents a truly bulk synchronous task graph with no
opportunity for task parallelism. For the stencil pattern, the
dominating factor is the basic overhead of executing a task, which is
minimal for MPI as the Task Bench implementation is bulk synchronous
and simply executes tasks one after another in alternation with
communication phases. In contrast, the asynchronous execution
mechanisms of the other systems are pure overhead in this scenario.

The gap between MPI and other systems shrinks as the complexity of the
communication pattern grows, and even reverses as task parallelism is
added in the form of multiple task graphs. Some of the slower systems
are omitted from comparisons with more complicated dependence
patterns, as they did not complete in reasonable time.

@@

The difference between the remaining systems is due primarily to the
scheduling behavior of the various systems. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Therefore, systems that provide an
additional on-node work stealing capability see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at very small task
granularities. This can be seen in the difference between the Chapel
distrib (work stealing) vs default (non-work-stealing) schedulers.

@@

Also of interest is the ability of asynchronous systems to
hide communication latency in the presence of task
parallelism. Figure~\ref{fig:efficiency-communication} plots efficiency vs task
granularity curves for varying amounts of
communication, determined by the number
of bytes produced by each task (and therefore communicated with each
task dependency).
